---
title: "Headlines"
author: "Audrey Smyczek"
date: "4/14/2022"
output: 
  html_document:
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)     # for graphing and data cleaning
library(lubridate)     # for date manipulation
library(ggthemes)      # for even more plotting themes
library(gganimate)     # for adding animation layers to ggplots
library(RColorBrewer)  # for color palettes
library(viridis)
library(plotly)        # for the ggplotly() - basic interactivity
library(gganimate)     # for adding animation layers to ggplots
library(transformr)    # for "tweening" (gganimate)
library(gifski)        # need the library for creating gifs but don't need to load each time
library(gt)
theme_set(theme_minimal()) # My favorite ggplot() theme :)
```





```{r}
freq_theme_words <- read.csv("https://raw.githubusercontent.com/the-pudding/data/master/women-in-headlines/word_themes_freq.csv")
freq_country_words <- read.csv("https://raw.githubusercontent.com/the-pudding/data/master/women-in-headlines/word_country_freq.csv")
sites <- read.csv("https://raw.githubusercontent.com/the-pudding/data/master/women-in-headlines/headlines_site.csv")
word_theme_rank <- read.csv("https://raw.githubusercontent.com/the-pudding/data/master/women-in-headlines/word_themes_rank.csv")
headlines <- read.csv("https://raw.githubusercontent.com/the-pudding/data/master/women-in-headlines/headlines.csv")
```

```{r}
polarity <- read.csv("https://raw.githubusercontent.com/the-pudding/data/master/women-in-headlines/polarity_comparison_site_country_time.csv")
```


```{r}
polarity_year <- read.csv("https://raw.githubusercontent.com/the-pudding/data/master/women-in-headlines/polarity_comparison_country_time.csv")
```


# headline examples

```{r}
headlines %>%
  select(site, headline_no_site, country, bias)%>%
  group_by(site)%>%
  filter(bias == max(bias) | bias == min(bias))%>%
  distinct(bias, .keep_all = TRUE)%>%
  gt()%>%
    tab_header(title = "Subset of Headlines with the Largest Bias Scores")
  
```


```{r}
# max table
headlines %>%
  select(site, headline_no_site, country, bias)%>%
  arrange(desc(bias))%>%
  distinct(site, .keep_all = TRUE)%>%
  slice(1:10) %>%
  gt()%>%
    tab_header(title = "Subset of Headlines with the Largest Bias Scores")
  
```

```{r}
# min table
headlines %>%
  select(site, headline_no_site, country, bias)%>%
  arrange(bias)%>%
  distinct(site, .keep_all = TRUE)%>%
  slice(1:10) %>%
  gt()%>%
    tab_header(title = "Subset of Headlines with the Smallest Bias Scores")
  
```

describe how bias is calculated 

# Polarity


```{r, fig.height= 5}
polarity %>%
  filter(country_of_pub == "UK")%>%
  ggplot()+
  geom_point(aes(x = polarity_base, y = fct_reorder(site, polarity_women)), color = "skyblue", size = 2)+
  geom_point(aes(x = polarity_women, y = fct_reorder(site, polarity_women)), color = "tomato2", size = 2)+
  geom_segment(aes(x = polarity_base, xend = polarity_women, y = site, yend = site))
```


```{r}
headlines%>%
  group_by(site)%>%
  filter(country == "USA")%>%
  summarize(mean_bias = mean(bias))%>%
  ggplot(aes(y = fct_reorder(site, mean_bias), x = mean_bias))+
  geom_point()

```

```{r}
headlines%>%
  filter(country == "USA")%>%
  ggplot(aes(y= site, x = bias))+
  geom_violin()
```

## polarity over time

```{r}
polarity_year%>%
  group_by(year)%>%
  summarize(mean_women = mean(women_polarity_mean), mean_all = mean(all_polarity_mean))%>%
  ggplot()+
    geom_smooth(aes(x=year, y= mean_women), se = FALSE, color ="skyblue")+
    geom_smooth(aes(x=year, y= mean_all), se = FALSE, color = "black")+ 
    #geom_point(aes(x = 2019, y = mean_women))+
    theme_classic()+
    labs(title = "Headline Polarity", y = "average polarity of headlines", x = NULL)+
    scale_x_binned()
```


A cumulative bar graph for the words used to describe women used in headlines. They are divided into 5 main categories with crime and violence having the most words and the highest frequency. The graph is interactive so each word can be highlighted with the individual word and frequency.

```{r}
pivot_words <- freq_theme_words %>% 
  pivot_longer(cols = -theme,
               names_to = "word",
               values_to = "freq") %>% 
  na.omit()

word_plot <- pivot_words %>% 
  filter(theme != "No theme") %>% 
  ggplot(aes(x = theme, 
             y = freq, 
             fill = fct_reorder(word, freq),
             text = paste("word:", word))) +
    geom_col(color = "black") +
    theme(legend.position = "none") +
    labs(title = "Cumulative Frequency of Words describing Women in Headlines",
       x = "",
       y = "Frequency")

ggplotly(word_plot,
         tooltip = c("y", "text"))
```


```{r}
pivot_country_word <- freq_country_words %>% 
  pivot_longer(cols = -country,
               names_to = "word",
               values_to = "number") %>% 
  filter(word != "X") %>% 
  na.omit()

pivot_country_word
```


```{r}

word_theme_rank %>% 
  filter(`rank` < 6) %>% 
  select(!`X`) %>% 
  ggplot(aes(y = word, x = count)) +
  geom_col(aes(fill = count))+
  scale_fill_viridis_c(option = "inferno") +
  facet_wrap(~theme,
            scales = "free")+
  theme(legend.position = "none")+
  labs(title = "Count of Top 5 words per Theme",
       y = "",
       x = "")
```


